---
title: "Example 3"
output: html_document
---

In addition to the YAML parameters shown in the first example, this letter displays some more parameters



* `yourref`, `yourmail`, `myref` common parts of the reference line (German: Bezugszeichenzeile)
* image in `signature` to make it look like it was hand signed.
* `signature-before` is for fine tuning the vertical space between closing and signature.
* `lang` sets the language of the document influences the labels of the reference line, date etc.. requires the respective latex language pack.
* `papersize` sets the paper size (default DIN A4).

<br>

### RMarkdown of Letter

```{r show-code, code=readLines('letter_example3.Rmd')[!grepl('ignette', x=readLines('letter_example3.Rmd'))], eval=FALSE, highlight=FALSE}

```

<br>

### Resulting PDF Letter

PDF rendered via `rmarkdown::render()` or the RStudio Knit button from RMarkdown document shown above.

```{r render-pdf, include=FALSE}
rmarkdown::render("letter_example3.Rmd", output_file="pkgdown_ex3.pdf", output_dir='.')

```

![Here should be a PDF of the letter](pkgdown_ex3.pdf){width=100% height=1000} \
