---
title: "Example 2"
output: html_document
---

In addition to the YAML parameters shown in the first example, this letter displays some more parameters

* `lco` via this parameter you can specify your own letter layout in a **L**etter **C**lass **O**ption file or use one of the layouts predefined by KomaScript (see *Introduction to komaletter*). Here KomaScript's US Letter Style is used.
* `papersize` to set a paper size other than DIN A4.
* `inline-links` causes links to be *hot* / active / clickable in the PDF and not implemented as footnote. You are still responsible for the styling of the linked text, here done as bold text.

<br>

### RMarkdown of Letter

```{r show-code, code=readLines('letter_example2.Rmd')[!grepl('ignette', x=readLines('letter_example2.Rmd'))], eval=FALSE, highlight=FALSE}

```

<br>

### Resulting PDF Letter

PDF rendered via `rmarkdown::render()` or the RStudio Knit button from RMarkdown document shown above.

```{r render-pdf, include=FALSE}
rmarkdown::render("letter_example2.Rmd", output_file="pkgdown_ex2.pdf", output_dir='.')

```

![Here should be a PDF of the letter](pkgdown_ex2.pdf){width=100% height=1000} \
