---
title: "Example 0"
description: "Template or Skeleton of a komaletter with a resonable small set of YAML parameters."
output: html_document
---

This example shows the skeleton you get when you run `rmarkdown::draft()` or start a letter from the template via the RStudio menu: File > New File > R Markdown... >  From Template > komaletter (PDF). 

It is intended as a starting point for your own letters and comes with a reasonably small set of settings in the YAML header. Actually, you could omit everything except `output: komaletter::komaletter' and still get a formatted letter. 

<br>

### Source Code of Letter

```{r show-code, code=readLines('letter_example0.Rmd')[!grepl('ignette', x=readLines('letter_example0.Rmd'))], eval=FALSE, highlight=FALSE}

```

<br>

### Resulting PDF Letter

PDF rendered via `rmarkdown::render()` or the RStudio Knit button from RMarkdown document shown above.

```{r render-pdf, include=FALSE}
rmarkdown::render("letter_example0.Rmd", output_file="pkgdown_ex0.pdf", output_dir='.')

```

![Here should be a PDF of the letter](pkgdown_ex0.pdf){width=100% height=1000} \
